Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs

### 2024-08-07 paper/manuscript.Rmd
Update manuscript fig + add guidelines for introduction review

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet (a), Loubna El Madouri (a, c), Maud Mouchet (a), Elisa Thébault (d) and Colin Fontaine (a)
date: Research report
output:
  #bookdown::word_document2:
  #  fontsize: 12pt
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: false
    fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
number_sections: true
bibliography: references.bib
csl: pnas.csl
---

@@

(c) ?

@@

@@

@@

@@

```{r sem, fig.cap = sem_cap}
knitr::include_graphics(here::here("figures", "fig2.pdf"))
```

@@

```{r}
std_cap <- paste0(
  "Direct and total effects of community biomass and species richness.
  Standardized coefficients ($r_\\delta$) derived from the temporal, spatial,
  and bioenergetic model Structural Equation Models. Points and lines display
  the estimated effects and their confidence intervals at 95\\%. Values are
  displayed in Table SXX, R-squared in Table SXX, and the visualisation of the
  bivariate relationships for the spatial and bioenergetic model SEMs are
  displayed in Figure SXX.")

@@

```{r std, fig.cap=std_cap, fig.width=7, fig.asp = .7}
knitr::include_graphics(here::here("figures", "p_sem_tps_sp.pdf"))
```

@@

Our results suggest that food-web that have undergone a biomass decline
experienced a loss of top trophic levels species and connectance, those gaining
biomass experienced a gain in top trophic levels and connectance. Importantly,
the same results were found using spatial comparison and theoretical models.
This results can be linked with the energetic and trophic biogeography theories
[@lindeman_trophic-dynamic_1942; @gravel_trophic_2011], which posits that the
presence of top predators is limited by the amount of energy and resource
available [@lindeman_trophic-dynamic_1942, @takimoto_ecosystem_2008;
@post_ecosystem_2000; @post_long_2002]. Thereby, a decrease in community biomass
can limit the amount of resource available to sustain top predators, while
increase in community biomass may result in more resource available for top
predators [@takimoto_ecosystem_2008; @post_ecosystem_2000].

@@

We simulated the food-webs dynamics across a gradient of global carrying capacity (K,
range: `r k_range["min"]`-`r k_range["max"]`). The carrying capacity were
standardized according to the number of producers and interspecific competition
($K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K$) to prevent changes in productivity
due to system size [@ives_stability_2000; @loreau_synchrony_2008]. We ran
simulations for 5000 timesteps. At the end of simulations, we removed dead
species and measured community biomass, species richness and food-web structure.
If the final community contained disconnected producers (i.e. without consumers)
or disconnected consumers (i.e. consumers without preys), we discarded them and
run the simulations again with the same parametrisation. The simulations were
done using the Julia package `EcologicalNetworksDynamics.jl`
[@lajaiiti_ecological_2024].

```{r matmet, fig.cap=fig_mat_cap, out.width="80%"}
knitr::include_graphics(get_mypath("figures", "mat_met.pdf"))
```

>>>>

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet (a), Loubna El Madouri (a, c), Maud Mouchet (a), Elisa Thébault (d) and Colin Fontaine (a)
date: "Research report"
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
always_allow_html: true
bibliography: references.bib
link-citations: true
#csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
csl: pnas.csl
---

@@

(c) Laboratoire d’Écologie Fonctionnelle et Environnement, UMR 5245
CNRS-UT3-INPT, Avenue de l’Agrobiopôle, Toulouse, France

@@

- Biodiversity changes = changes in food-web structure that are important for
  ecosystem functioning and stability.
- Unique combination of empirical data and theoretical simulations.
- Inform how community biomass and species richness declines result in an
  increase of the fragility of food-webs.

@@

- Ideas:
  1. Biodiversity at risk; widespread heterogeneity in changes; ignores
     community struture
  2. Community structure is important for ecosystem functioning and robustness
     to perturbations
  3. Questions: assembly and disassembly processes of food-webs associated with
     biodiversity changes?; possible due to inference of food-web structure;
     Energetic constraints, redundancy and robustness; Validity of space for
     time hypothesis
  4. Presentation of the study data, theoretical models and extinction
     simulation: presentation of the hypothesis associated to the SEM

Why redundancy? Because not all links in the same way to the robustness of
food-webs

@@

```{r matmet, fig.cap=fig_mat_cap, out.width="100%"}
knitr::include_graphics(get_mypath("figures", "mat_met.png"))
```

@@

```{r sem, fig.cap = sem_cap, out.width = "100%"}
tar_load(p_sem_tps_file)
knitr::include_graphics(p_sem_tps_file)
```

@@

```{r}
std_cap <- paste0(
  "Direct and total effects of community biomass and species richness on
  food-web structure. Standardized coefficients ($r_\\delta$) derived from the
  temporal, spatial, and bioenergetic model Structural Equation Models. Points
  and lines display the estimated effects and their confidence intervals at
  95\\%. Values are displayed in Table SXX, R-squared in Table SXX, and the
  visualisation of the bivariate relationships for the spatial and bioenergetic
  model SEMs are displayed in Figure SXX.")

@@

```{r std, fig.cap=std_cap, out.width = "100%"}
tar_load(p_semeff_tot_ok_file)
knitr::include_graphics(p_semeff_tot_ok_file)
```

@@

Our results suggest that food-web that have undergone a biomass decline
experienced a loss of top trophic levels species and connectance, those gaining
biomass experienced a gain in top trophic levels and connectance. Importantly,
the same results were found using spatial comparison and theoretical models.
This results can be linked with the energetic and trophic biogeography theories
[@lindeman_trophic-dynamic_1942; @gravel_trophic_2011], which posits that the
presence of top predators is limited by the amount of energy and resource
available [@lindeman_trophic-dynamic_1942, @takimoto_ecosystem_2008;
@post_ecosystem_2000; @post_long_2002]. Thereby, a decrease in community biomass
can limit the amount of resource available to sustain top predators, while
increase in community biomass may result in more resource available for top
predators [@takimoto_ecosystem_2008; @post_ecosystem_2000]. (Add something about
the role of apex predators?)

@@

We simulated the food-webs dynamics across a gradient of global carrying capacity (K,
range: `r k_range["min"]`-`r k_range["max"]`). The carrying capacity were
standardized according to the number of producers and interspecific competition
($K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K$) to prevent changes in productivity
due to system size [@ives_stability_2000; @loreau_species_2008]. We ran
simulations for 5000 timesteps. At the end of simulations, we removed dead
species and measured community biomass, species richness and food-web structure.
If the final community contained disconnected producers (i.e. without consumers)
or disconnected consumers (i.e. consumers without preys), we discarded them and
run the simulations again with the same parametrisation. The simulations were
done using the Julia package `EcologicalNetworksDynamics.jl`
[@lajaaiti_ecologicalnetworksdynamicsjl_2024].
